---
title: "adelaide_graph"
author: "Adelaide Robinson"
date: "2023-03-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(googlesheets4)
library(tidyverse)
```


Visualize some aspect of your data assignment for EDS-240.

#Read in the Data 
```{r}
url <- "https://docs.google.com/spreadsheets/d/1cUz4WZ1CRHFicuVUt82kJ_mL9Ur861Dn1c0BYu3NmRY/edit#gid=0"

MPS_tracker_data <- read_sheet(url)

test <- MPS_tracker_data |> group_by(year, site, sub_category) |> summarize(n()) 
```

#Wrangle Data

```{r}
new_data <- MPS_tracker_data |>
  filter(str_detect(site, "Galapagos")) |> 
  filter(! indicator_type == "Impact Indicators") |>  #remove Impact indicators 
  mutate(site = "Galapagos EEZ", score = as.numeric(score)) |>  #call it same site to connect EEZ with later data
 group_by(site, year, category) |>
  summarize(score = round(mean(score, na.rm = TRUE),0)) |> ungroup()  #find mean for each category 
  
# Stacked
ggplot(new_data, aes(fill=category, y=score, x=year)) + 
    geom_bar(position="stack", stat="identity") + labs(title = "Galapagos EEZ", y = "Average Score", x = "Year") + theme(text = element_text(size=20)) +  scale_fill_manual(values=c("#eb6c04",
                             "#00adaa",
                             "#003e70",
                             "#f68e8a",
                             "#6aaf5a"))

```

```{r}
  filter(str_detect(site, "Galapagos")) |> 
  filter(! indicator_type == "Impact Indicators") |>  #remove Impact indicators 
  mutate(site = "Galapagos EEZ", score = as.numeric(score)) |>  #call it same site to connect EEZ with later data
 group_by(site, year, category) |>
  summarize(score = round(mean(score, na.rm = TRUE),0))  #find mean for each category 

+ labs(title = "Galapagos EEZ", y = "Average Score", x = "Category", axis.title=element_text(size=14,face="bold")) + scale_fill_manual(values=c("#eb6c04",
                             "#00adaa",
                             "#003e70",
                             "#f68e8a",
                             "#6aaf5a"))
```


Doing the same thing but instead using total score. Need to make sure we are including the exact same categories each year, otherwise we won't be able to compare between years. 

```{r}
total_data <- MPS_tracker_data |>
  filter(str_detect(site, "Galapagos")) |> 
  filter(!indicator_type == "Impact Indicators") |> 
  mutate(score = as.numeric(score)) #we know impact indicators were not included 
  

#come up with the sub-categories that are in all years
dat_19 <- total_data |> filter(year == 2019) 

dat_20 <- total_data |> filter(year == 2020)
dat_21 <- total_data |> filter(year == 2021)
dat_22 <- total_data |> filter(year == 2022)

#inner join to keep only categories in both years
combined <- inner_join(dat_19, dat_20, by = "sub_category") |> select(sub_category)

#inner join to keep only categories in both years
combined_2 <- inner_join(combined, dat_22) |> select(sub_category)

#inner join to keep only categories in both years
new <- inner_join(combined_2, dat_21) |> filter(!sub_category == "Cost-Efficiency")

#list of categories in both years
list <- new$sub_category

final_test <- total_data |> filter(sub_category %in% list)


final_new <- final_test |> group_by(category, year) |> 
  summarise(total_score = sum(score), n())

# Stacked
ggplot(final_new, aes(fill=category, y=total_score, x=year)) + 
    geom_bar(position="stack", stat="identity") + labs(title = "Galapagos EEZ", y = "Total Score") + scale_fill_manual(values=c("#eb6c04",
                             "#00adaa",
                             "#003e70",
                             "#f68e8a",
                             "#6aaf5a"))

```

Line Graph

```{r}
new_data <- MPS_tracker_data |>
  filter(str_detect(site, "Yuca")) |> 
  filter(! indicator_type == "Impact Indicators") |>  #remove Impact indicators 
  mutate(site = "Gabon EEZ", score = as.numeric(score)) |>  #call it same site to connect EEZ with later data
 group_by(site, year, category) |>
  summarize(score = round(mean(score, na.rm = TRUE),0)) |> ungroup()  #find mean for each category 


new_data |> ggplot(aes(y = score, x = year, group = 1)) + geom_line(aes(col = category, group = category), position=position_dodge(w=.2), size =1) + ylim(c(1,5)) + labs(title = "Score by Category at Yucatán, Mexico", y= "Average Score")  + scale_color_manual(values=c("#eb6c04",
                             "#00adaa",
                             "#003e70",
                             "#f68e8a",
                             "#6aaf5a")) + theme(
                             plot.title = element_text(size =20), axis.title = element_text(size = 15), legend.text = element_text(size = 15),legend.title = element_text(size = 15),
                             
         panel.background = element_rect(fill='transparent'),
         plot.background = element_rect(fill='transparent', color=NA),
         panel.grid.major = element_line(color = "grey"),
         panel.grid.minor = element_blank(),
         legend.background = element_rect(fill='transparent'),
         legend.box.background = element_rect(fill='transparent')
       )
                             

new_data |> ggplot(aes(y = score, x = year, group = 1)) + geom_line(aes(col= category), size = 2) + ylim(c(1,5)) + labs(title = "Score by Category at Yucatán, Mexico", y= "Average Score")  + scale_color_manual(values=c("#eb6c04",
                             "#00adaa",
                             "#003e70",
                             "#f68e8a",
                             "#6aaf5a")) + facet_wrap(~category) + theme(
         panel.background = element_rect(fill='transparent'),
         plot.background = element_rect(fill='transparent', color=NA),
         panel.grid.major = element_line(color = "grey"),
         panel.grid.minor = element_blank(),
         legend.background = element_rect(fill='transparent'),
         legend.box.background = element_rect(fill='transparent')
       )
                               