I think a ridgeline plot could be really cool... see this website on hoe to make one: https://r-graph-gallery.com/294-basic-ridgeline-plot.html
I got some ideas from the website "from data to viz"

For the map, I think it would be really cool to show a density plot of the total scores for that site
this would be density on the y axis and score 1-5 on the x axis 
here's a website on how to make a density plot:
https://r-graph-gallery.com/21-distribution-plot-using-ggplot2.html

Load libraries 
```{r}
library(ggplot2)
library(ggridges) 
library(googledrive) 
library(googlesheets4) 
library(dplyr)
library(tidyverse)
```


Read in the data 
```{r}
url <- "https://docs.google.com/spreadsheets/d/1cUz4WZ1CRHFicuVUt82kJ_mL9Ur861Dn1c0BYu3NmRY/edit#gid=0"

MPS_tracker_data <- read_sheet(url)
```

Practice making a ridgeline plot with the diamonds dataset 
```{r}
ggplot(diamonds, aes(x = price, y = cut, fill = cut)) +
  geom_density_ridges() +
  theme_ridges() + 
  theme(legend.position = "none")
```

Now let's try that with our data 
```{r}
# let's try to put the country on the y axis and the average score on the x axis 

# make a mini data frame of what we want 
# we have the problem that if we are to compare TOTAL scores accross countries, some countries have more sites and observations... this is why I'm going with the average for now
# ACTUALLY uh oh... you can't do this type of plot with the grouped averages 

# # this object is useless right now: 
# country_total_scores <- MPS_tracker_data |> 
#   mutate(as.factor(country)) |> 
#   group_by(country) |> 
#   summarise(sum_score = mean(as.numeric(score), na.rm = TRUE), na.rm = TRUE) |> 
#   na.omit() # how do I get rid of NAs?

# take out all rows that have "don't know or NA"
country_scores_no_na <- MPS_tracker_data[!(MPS_tracker_data$score == "Don't know or N/A"), ] |> 
  na.omit() |> 
  select(country, score)

  
ggplot(country_scores_no_na, aes(x = score, y = country, fill = country)) +
  geom_density_ridges() +
  theme_ridges() 
```

```{r}
# another tester one: 
# library
library(ggridges)
library(ggplot2)
library(viridis)
library(hrbrthemes)

# Plot
ggplot(lincoln_weather, aes(x = `Mean Temperature [F]`, y = `Month`, fill = ..x..)) +
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01) +
  scale_fill_viridis(name = "Temp. [F]", option = "C") +
  labs(title = 'Temperatures in Lincoln NE in 2016') +
  theme_ipsum() +
    theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8)
    )
```

