I think a ridgeline plot could be really cool... see this website on hoe to make one: https://r-graph-gallery.com/294-basic-ridgeline-plot.html
I got some ideas from the website "from data to viz"

For the map, I think it would be really cool to show a density plot of the total scores for that site
this would be density on the y axis and score 1-5 on the x axis 
here's a website on how to make a density plot:
https://r-graph-gallery.com/21-distribution-plot-using-ggplot2.html

Load libraries 
```{r}
library(ggplot2)
library(ggridges) 
library(googledrive) 
library(googlesheets4) 
library(dplyr)
library(tidyverse)
```


Read in the data 
```{r}
url <- "https://docs.google.com/spreadsheets/d/1cUz4WZ1CRHFicuVUt82kJ_mL9Ur861Dn1c0BYu3NmRY/edit#gid=0"

MPS_tracker_data <- read_sheet(url)
```

Practice making a ridgeline plot with the diamonds dataset 
```{r}
ggplot(diamonds, aes(x = price, y = cut, fill = cut)) +
  geom_density_ridges() +
  theme_ridges() + 
  theme(legend.position = "none")
```

Now let's try that with our data 
```{r}
# let's try to put the country on the y axis and the average score on the x axis 

# make a mini data frame of what we want 
# we have the problem that if we are to compare TOTAL scores accross countries, some countries have more sites and observations... this is why I'm going with the average for now
# ACTUALLY uh oh... you can't do this type of plot with the grouped averages 

# # this object is useless right now: 
# country_total_scores <- MPS_tracker_data |> 
#   mutate(as.factor(country)) |> 
#   group_by(country) |> 
#   summarise(sum_score = mean(as.numeric(score), na.rm = TRUE), na.rm = TRUE) |> 
#   na.omit() # how do I get rid of NAs?

# take out all rows that have "don't know or NA"
country_scores_no_na <- MPS_tracker_data[!(MPS_tracker_data$score == "Don't know or N/A"), ] |> 
  na.omit() |> 
  select(country, score)

  
ggplot(country_scores_no_na, aes(x = score, y = country, fill = country)) +
  geom_density_ridges(scale = 1.2) +
  theme_ridges() 
```

```{r}
# another tester one: 
# library
library(ggridges)
library(ggplot2)
library(viridis)
library(hrbrthemes)

# Plot
ggplot(lincoln_weather, aes(x = `Mean Temperature [F]`, y = `Month`, fill = ..x..)) +
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01) +
  scale_fill_viridis(name = "Temp. [F]", option = "C") +
  labs(title = 'Temperatures in Lincoln NE in 2016') +
  theme_ipsum() +
    theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8)
    )
```

Try making a density plot comparing two things with transparency
```{r}
# compare surveillance and enforcement scores across all sites in 2022
# by the way I think I deleted "MPS_no_na" somehow so you'll have to recreate it later
MPS_no_na <- MPS_tracker_data |> na.omit()

surveillance_df <- MPS_no_na |> 
  filter(category == "Surveilance and Enforcement", 
         site == c("Bahamas EEZ", 
                   "Celestun Fishery Refuge", 
                   "Costa del Ecuador", 
                   "National Parcs", 
                   "PAN", 
                   "Scorpion Reef National Park (Arrecife Alacrane Bioshpere Reserve)"))

  

ggplot(data = surveillance_df, 
       aes(x = score, group = site, fill = site)) + 
  geom_density(adjust = 1.5, alpha = 0.5) + 
  theme(legend.position = "bottom") + 
  theme_ipsum() + 
  labs(title = "Density of surveillance and enforcement scores across five sites")
```

# I have an idea! compare certain aspects of the enforcement strategy with the END GOAL. There are a couple end goals that we could go for: 
- decreased bicatch 
- community benefits 
- decreased protected hunting 
- declining infractions 
- marine population status
possibly more! 

a useful question to ask could be something like: how does community involvement affect decreased bicatch... haha that could be interesting to understand! BUT we would need a different type of dataset for this... I don't even think that this dataset is tidy yet!! we need to have a single row for an entire year at a certain site... how can we do this? pivot _wide() ??
obviously there is not much data for making specific inferences yet, but this program provides a framework to eventually do that... and we should brag about the fact that this is the first tool to track and monitor the marine enforcement efforts. This is why this is so important... say stuff like this in the next presentation!! 




